Incorporating periodic variability in hidden Markov models for animal movement

### 2016-07-27 paper3.Rnw
pictures with nice legends, paragraph shortly

sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev)
   + scale_colour_brewer(palette="Dark2")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

@@

mnames <- c("fmm5","fmmsin4","hmm6","hmmblock4","hmmhourly3",
            "hmmquad5","hmmsin5","obs")
mnames2 <- c("fmm-homogeneous  (n=5)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=6)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=5)"
             , "sin (n=5)"
             , "observed")

combdf <- (rbind(avgdf,obsdf)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6)),1)
colvec <- c(c(brewer.pal(7,"Dark2"),"black"),"black")
lwdvec <- c(rep(c(1,4),c(7,1)),4)
alphavec <- c(rep(c(1,0.2),c(7,1)),1)
(ggplot(combdf,aes(x=time,y=steplen,colour=model,lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Time of Day") 
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitdf,aes(y=viterbi),
               lwd=2.5,alpha=1,colour="black",shape=118,show.legend=FALSE)
## adding geom_point() with vitdf screws up the legend ...
@

@@

acfdf <- (readRDS("summary_stats/cat.1.df.RDS")$acfdf 
          %>% mutate(model = .id)
)
(ggplot(acfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_colour_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitacf,lwd=2.5,alpha=1,colour="black",shape=118,
             )
## vitacf messes up legend ...
@

>>>>

sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

@@

vitdf2 <- vitdf %>% transmute(time=time,variable="viterbi",value=viterbi)

mnames <- c("fmm5","fmmsin4","hmm6","hmmblock4","hmmhourly3",
            "hmmquad5","hmmsin5","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=5)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=6)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=5)"
             , "sin (n=5)"
             , "observed"
             , "viterbi homogenous (n=3)")

combdf <- (rbind(avgdf,obsdf,vitdf2)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6))
                ,0
               )
colvec <- c(c(brewer.pal(7,"Dark2"),"black")
             ,"black"
            )
lwdvec <- c(rep(c(1,4),c(7,1))
             ,4
            )
alphavec <- c(rep(c(1,0.2),c(7,1))
               ,1
              )

shapevec <- c(rep(32,8),118)

(ggplot(combdf,aes(x=time,y=steplen,colour=model,
                   lwd=model,
                   alpha=model,linetype=model, shape=model))
  + geom_point()
  + geom_line()
  + xlab("Time of Day") 
  + scale_shape_manual(values=shapevec, name="Model")
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
 ) 
# + geom_point(data=vitdf,aes(y=viterbi),
               # lwd=2.5,alpha=1,colour="black",shape=118,show.legend=FALSE)
## adding geom_point() with vitdf screws up the legend ...
@

@@

acfdf <- (readRDS("summary_stats/cat.1.df.RDS")$acfdf 
          %>% mutate(model = .id)
)

combacfdf <- (rbind(acfdf,vitacf)
           %>% transmute(lag=lag,model=model,ACF=ACF)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)

(ggplot(combacfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model,shape=model))
  + geom_point()
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_shape_manual(values=shapevec, name="Model")
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
)

## vitacf messes up legend ...
@
